---
title: "SMH R Workshop - Data Manipulation"
output: html_document
---

Often times, data does not come the way you want it for analysis. Data manipulation and processing are key steps in data analysis, and can often take the longest.

The dplyr package is part of the tidyverse, and is a great package that will serve the majority of your data processing needs. We will be going over 4 of the most commonly used functions: 1) filter()/select(), 2) mutate(), 3) summarise(), and 4) group_by().


# Installing and loading packages

First, we need to install and load the packages that we will be using. We will use the readr package for importing data into R, and the dplyr package for data processing.

```{r}
# Using install.packages(), install the dplyr and readr packages

install.packages("dplyr")
install.packages("readr")
# Now, load these two packages

library(dplyr)
library(readr)

```

# Importing data

In this project folder, I've included a dataset named winequality-red.csv. Let's import this and explore it.

```{r}

# Using read_csv() from readr to import the dataset
wine_data <- read_csv("winequality-red.csv")

# Use the summary() function to get a quick summary of the dataset
summary(wine_data)

```


# Filtering rows and selecting columns

```{r}

# The filter() and select() columns serve two of the most basic data processing steps: filtering down your rows, and selecting the right columns

# Let's create a dataset that contains only acidic wines by filtering wines that have a fixed.acidity of less than 7

acidic_wines <- filter(wine_data, fixed.acidity < 7)

summary(acidic_wines) # Notice the range of the fixed.acidity value

# Let's now focus on the density, pH, and sulphates level. Create a data subset that contains only these three columns

acidic_wines_subset <- select(acidic_wines, density, pH, sulphates)

summary(acidic_wines_subset)

```

# The pipe operator

One of the great things about working in the tidyverse is the ability to "pipe" operations together using the pipe operator ("%>%"). Piping allows you to combine related operations and weave them together to tell a complete story.

```{r}
# From the original wines_data dataset, create a new dataset named new_data that contains only acidic wines (fixed.acidity < 7) and the columns density, pH, and sulphates

new_data <- wine_data %>%
  filter(fixed.acidity < 7) %>%
  select(density, pH, sulphates)

summary(new_data)

```

# Creating new variables using mutate()

```{r}

# Create a new variable named acidity_level that is the "low" if fixed.acidity is less than 10, and "high" otherwise
# We can use the builtin ifelse() function for conditional assignments

wine_data <- wine_data %>%
  mutate(acidity_level = ifelse(fixed.acidity < 10, "low", "high"))

head(wine_data)

```

# Creating data summaries

The summarise() function is a versatile function that allows you to do a lot of different things. Generally, you use this function when you want to summarise/aggregate your date

```{r}
# Using summarise(), calculate the mean, median, minimum, and maximum of the pH of all wines in the wine_data dataset

wine_data %>%
  summarise(mean_pH = mean(pH),
            median_pH = median(pH),
            min_pH = min(pH),
            max_pH = max(pH))


```

# Grouping operations using group_by()

Sometimes, grouped operations can be useful when you want to perform different calculations on subgroups of your data. The group_by() function from dplyr allows you to group sets of operations together. 

```{r}
# Using the wine_data dataset and the acidity_level variable we created, calculate the mean, SD, minimum, and maximum pH levels of all wines, grouped by the acidity_level variable

wine_data %>%
  group_by(acidity_level) %>% 
  summarise(mean_pH = mean(pH),
            sd_pH = sd(pH),
            min_pH = min(pH),
            max_pH = max(pH))

```

