---
title: "SMH R Workshop - Data Visualization"
output: html_document
---

So far, you've been using functions included in base R, but there are many more R packages that contain additional useful functions for various tasks. One such package is ggplot2, which is a very popular data visualization package.


# Packages

## Installing packages

```{r}
# To install R packages (works for most packages)

install.packages("package_name") # Note that the package_name is wrapped in double quotes. This is important.

# Try installing the ggplot2 package

install.packages("ggplot2")
```

You only need to install packages once on any given computer.

## Loading packages

Once you've installed a package, in order to access the package's functions, you would need to load the package. In contrast to installation, this needs to be done every time you open and re-open R.

```{r}
# To load an R package, use the library() function

# Try loading the ggplot2 package

library(ggplot2)

# Once you've loaded the package, you can access the package contents

```

