---
title: "SMH R Workshop - R basics"
output: html_document
---

# R and RStudio Basics

This is an RMarkdown document. RMarkdown can be used to create a variety of outputs (e.g. reports), but for now we will be using it as a way to document our data analysis. 

You can type text surrounding code chunks, as you would in a word document.

```{r}

# This is a code chunk. Within code chunks, you can type R code to execute operations. To create a new code chunk, you can use the "Insert Chunk" button in the toolbar, or press "Ctrl/Cmd + Alt + I" on your keyboard.

# When typing R code, you may also wish to insert comments to annotate your code (like this one!). You can do this simply by starting a line with the "#" sign

# To execute code from a code chunk, either select the lines of code and press "Run" in the toolbar, or press "Ctrl/Cmd + Enter" on your keyboard



# Try running the following lines of code
1 + 1
(1 + 3) / 2
1 + 3 / 2

```

# R Programming Basics

## R Objects

```{r}

# Everything in R is stored in R objects

# All objects are created in some form of the following

object_name <- value

# E.g. Assign 'x' the value of 2
x <- 2
x

# The '<-' is R's assignment operator. The direction of the arrow signifies what is the object name and 
# what is the value being assigned to the object.

# Note: Functionally, the '=' operator is the same as '<-'. However, it is highly recommended that you use the '<-' operator 
#       to avoid confusion later down the road as you write more complex code.

# Don't do this!
y = 3

# Instead, do this
y <- 3

```

R object names can only contain letters, numbers, underscores, and periods. They must start with a letter. Using descriptive object names are highly recommended. There are several coding conventions for object names with multiple words.

```{r}

# Snake case - the preferred one for most people

snake_case_is_great

# Camel case - another commonly used one

camelCaseIsOkay

# Periods

using.periods.is.not.as.good

# Please don't do this

what_do.iEven_call.this

```

## Using functions

R has a lot of built-in functions, and there are many more functions that can be loaded additionally through external packages. Functions allow users to do more operations.

```{r}

# R functions are called like this:

function_name(arg1, arg2, ...)

# A function call takes in arguments (some required, some optional) and does computations based on the inputted parameters

# E.g. the sum() function allows users to calculate a sum
x <- 2
y <- 5
sum(x, y)

```

## help() function

Brief sidenote: R has a built-in help function that allows you to pull up the helpfile for any function you have access to (including base R functions, as well as any functions from externally loaded packages). Helpfiles are very useful as you troubleshoot your code.

```{r}
# To access the helpfile for the sum() function, you can use
help(sum)

# OR, simply use
?sum
```



## Basic troubleshooting examples

Now that you're familiar with some R basics, see if you can complete the missing code!

```{r}

# Assign 29 to an object named x1
x1 <- 29
x1
# Create another object named x2, and assign it the value of 34
x2 <- 34

# Using the sum() function, calculate the sum of x1 and x2, and save this to an object named total
total <- sum(x1, x2)
total 
# Check that the value of total is 63
print(total)  

```


# Next section: ggplot2 slides