---
title: "SMH R Workshop - Data Visualization"
output: html_document
---

So far, you've been using functions included in base R, but there are many more R packages that contain additional useful functions for various tasks. One such package is ggplot2, which is a very popular data visualization package.


# Packages

## Installing ggplot2 package

```{r}
# To install R packages (works for most packages)

install.packages("package_name") # Note that the package_name is wrapped in double quotes. This is important.

# Try installing the ggplot2 package

install.packages("ggplot2")
```

You only need to install packages once on any given computer.

## Loading ggplot2

Once you've installed the ggplot2 package, in order to access the package's functions, you would need to load the package. In contrast to installation, this needs to be done every time you open and re-open R.

```{r}
# To load an R package, use the library() function

# Try loading the ggplot2 package

library(ggplot2)

# Once you've loaded the package, you can now access the package contents

```


# The iris dataset

For the following data visualization exercises, we will use the built-in "iris" dataset. 

```{r}
# Load the data
data(iris)

# Brief summary of the data
summary(iris)

```

# Data visualization with ggplot2

## Visualizing univariate distributions

```{r}
# Univariate continuous distributions can be visualized in various ways, but one popular way is a histogram

# Let's create a basic histogram of the sepal length variable from the iris dataset using ggplot2
ggplot(data = iris,
       aes(x = Sepal.Length)) + # Specify which variable to map to the x-axis
  geom_histogram() # Add a geom_histogram layer to create a histogram of sepal length


# Try adapting the previous code to create a histogram of the sepal width





```

##
```{r}

```

