---
title: "SMH R Workshop - Data Manipulation"
output: html_document
---

Often times, data does not come the way you want it for analysis. Data manipulation and processing are key steps in data analysis, and can often take the longest.

The dplyr package is part of the tidyverse, and is a great package that will serve the majority of your data processing needs. We will be going over 4 of the most commonly used functions: 1) filter()/select(), 2) mutate(), 3) summarise(), and 4) group_by().


# Installing and loading packages

First, we need to install and load the packages that we will be using. We will use the readr package for importing data into R, and the dplyr package for data processing.

```{r}
# Using install.packages(), install the dplyr and readr packages

install.packages("dplyr")
install.packages("readr")
# Now, load these two packages

library(dplyr)
library(readr)

```

# Importing data

In this project folder, I've included a dataset named winequality-red.csv. Let's import this and explore it.

```{r}

wine_data <- read_csv("winequality-red.csv")


```


# Filtering rows and selecting columns

```{r}
# The filter() and select() columns serve two of the most basic data processing steps: filtering down your rows, and selecting the right columns



```


